---
output: github_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "fig/README-"
)

library(covr)
coverage_detail <- package_coverage()
coverage <- percent_coverage(coverage_detail) |>
  round(0)

cov_col <- cut(
  coverage,
  breaks = c(-Inf, 25, 50, 70, 85, 90, 100),
  c("red", "orange", "yellow", "yellowgreen", "green", "brightgreen")
)
```

<!-- badges: start -->
[![R-CMD-check](https://github.com/animalnexus/feedr/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/animalnexus/feedr/actions/workflows/R-CMD-check.yaml)
![Code Coverage: `r coverage`%](https://img.shields.io/badge/code_coverage-`r coverage`%-`r cov_col`)
<!-- badges: end -->
  

# feedr

feedr is an R package designed to load, transform and visualize RFID data created when pit-tagged animals are detected by RFID loggers associated with static stations (e.g., seed-feeders, nest boxes, nectar-feeders, etc.).

## Installation

To ensure the smoothest possible installation, it is recommended that you start a new R session before installing. 

feedr is available from GitHub and can be installed directly with the package 'pak'. 

```{r}
#| eval: false
install.packages("pak") # if not already installed
pak::pkg_install("animalnexus/feedr")
```

## History & Citation

feedr was originally (back in 2015/2016) developed as an R package *and* Shiny
App to be hosted by the animal**nexus** project. This was a large collaborative effort to develop tools for the observation, visualization, and analysis of animal movements registered by RFID feeders or other static recording stations.

However, over the years the group has been exploring other avenues of research. 

In order to keep maintenance of the feedr package manageable, it was split into two packages, the core functions (feedr) and the Shiny UI ([feedrUI](https://github.com/animalnexus/feedrUI)).

**The feedr package is actively maintained by Steffi LaZerte, 
but the feedrUI package is currently on hold.**

We published an introduction to the package in Ecology in Evolution in 2017 which you
can use to cite this package:

LaZerte SE, Reudink MW, Otter KA, Kusack J, Bailey JM, Woolverto, A, Paetkau M, de Jong A, and Hill DJ. feedr and animalnexus.ca: A paired R package and user-friendly Web application for transforming and visualizing animal movement data from static stations. Ecol Evol. 2017; 7: 7884–7896. https://doi.org/10.1002/ece3.3240 


## Demos
There is a demo for using `map_leaflet()` which you can run as follows:

```{r}
#| eval: false
demo("demo-leaflet", package = "feedr")
```
