---
title: "Conversion functions"
vignette: >
  %\VignetteIndexEntry{Conversion functions}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
knitr:
  opts_chunk:
    collapse: true
    comment: '#>'
---

Conversion functions permit you to format raw or transformed data for use in other packages.

## Setup

First we'll load the packages we want to use.

```{r}
#| message: false
library(feedr)
```

### `convert_anidom()`

Converts the output of `disp()` for use by the package [`aniDom`](https://cran.r-project.org/package=aniDom)'s functions: `elo_scores()` and `estimate_uncertainty_by_splitting()`

```{r}
d <- disp(visits(finches_lg), bw = 5)
i <- convert_anidom(d)
i
```

```{r}
library(aniDom)

# Calculate elo_scores
s <- elo_scores(winners = i$winner, losers = i$loser)
s

# Estimate repeatability
estimate_uncertainty_by_repeatability(winners = i$winner, losers = i$loser)
estimate_uncertainty_by_splitting(
  winners = i$winner,
  losers = i$loser,
  randomise = TRUE
)
```

```{r eval = FALSE}
# Construct social network graphs
Sociogram(
  data_sheet = i$data_sheet,
  items = i$items,
  actions = i$actions,
  bits = i$bytes
)
```

### `convert_perc()`
Converts the output of `disp()` for use by the package [`Perc`](https://cran.r-project.org/package=Perc)

```{r}
d <- disp(visits(finches_lg), bw = 5)
i <- convert_perc(d)
head(i)
```

```{r}
library(Perc)

# Calculate ranks (adapted from Perc examples)
conflict_mat <- as.conflictmat(i, weighted = TRUE)
perm <- conductance(conflict_mat, 2)
simRankOrder(perm$p.hat, num = 10, kmax = 1000)
```

### `convert_asnipe()`
Converts raw RFID data into a format for easy use by either the `gmmevents()` or the `get_associations_points_tw()` functions included in the [`asnipe`](https://cran.r-project.org/package=asnipe) package for calculating group membership.

For `gmmevents()`:
```{r}
a <- convert_asnipe(finches, fun = "gmmevents")
head(a)
```

```{r, eval = FALSE}
library(asnipe)
gmmevents(a$time, a$identity, a$location)
```

For `get_associations_points_bw()`:
```{r}
a <- convert_asnipe(finches, fun = "get_associations_points_tw")
head(a)
```

```{r, eval = FALSE}
get_associations_points_tw(a)
```


### `convert_activity()`
Converts raw RFID data into a format for easy use by the `fitact()` function included in the [`activity`](https://cran.r-project.org/package=activity) package for modelling activity levels and daily activity patterns.

```{r}
i <- convert_activity(finches_lg)
```

```{r}
library(activity)

# Calculate daily activity pattern for a single individual
a <- fitact(i[[1]], sample = "none")
plot(a)

# Calculate daily activity pattern for all individuals
a <- lapply(i, fitact, sample = "none")
plot(a[[3]])
plot(a[["06200004F8"]])
```
