---
title: "Housekeeping"
author: "Steffi LaZerte"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Housekeeping}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<a id = "top"></a>

```{r, include = F}
library(feedr)
```

This is a quick tutorial covering a couple functions you can use to clean and fix problems in your data.

- [`check.ids()`](#checkids)
- [`check.problems()`](#checkproblems)

These functions can be skipped, or used one after the other. It's up to you. Basically, they are wrapper functions for cleaning your data while maintaining appropriate categorical (factor) levels.

This tutorial assumes that you already know how to [load/import](load.html) your data. We will be using the feedr package example dataset: finches.

```{r}
head(finches)
```

<a id = "checkids"></a>

## `check.ids()`
This function can be used to remove any bird\_ids that are present, but which you know aren't really birds. For example, if you use a 'wand' to test the deployment of your feeders, this is a bird\_id that you should remove prior to analysis. Further, occasionally there are bird\_ids that are error codes (e.g. 0000000000), you may wish to determine why these are present, but once again, for analysis they should be removed.

This function works by comparing the list of bird\_ids in the data to an external, bird\_id dataset. The dataset is expected to have at least two columns: bird\_id and species. The species column should either contain species identity (e.g. House Finch) or the error code (e.g. wand, or error).

In addition to removing error or wand ids, this function will also report which bird\_ids are in your data sets, but not in your bird\_id index, and which are in your bird\_id index, but not in your data sets.

### With no error/wand ids

Let's load a bird index file. Note that there are no errors or wands ids.

```{r eval = FALSE}
bird_index <- read.csv("./Data/bird_index.csv")
bird_index
```

```{r echo = FALSE}
bird_index <- read.csv(system.file("extdata", "bird_index.csv", package = "feedr"))
bird_index
```

You need to give this function a data set, and it will return a cleaned data set. Here we'll save it as r.clean:
```{r}
r.clean <- check.ids(finches, bird_ids = bird_index)
```

This output shows that all the bird\_ids in the data are also in the index and vice versa. Further, there were no omitted ids (error or wand ids).

### With error/wand ids
Let's see how it works if you did have a 'wand' or 'error' code in your index file that matched a bird\_id in your dataset.

```{r echo = 4}
bird_index$species <- as.character(bird_index$species)
bird_index$species[1] <- "wand"
bird_index$species[2] <- "error"
bird_index
```

```{r}
r.clean <- check.ids(finches, bird_ids = bird_index)
```

Here we omited the two bird\_ids associated with a wand (0620000514) and with an error (041868D861).  

Note that nothing else changed.

### bird\_ids present in dataset but not in the index
```{r echo = 2}
bird_index <- bird_index[-5, ]
bird_index
```

```{r}
r.clean <- check.ids(finches, bird_ids = bird_index)
```


<a id = "checkproblems"></a>

## `check.problems()`

This function is only necessary if for some reason, you're getting errors in the bird\_ids in your data set.

This function will correct all instances of a bird\_id according to the list provided.

```{r eval = FALSE}
problems <- read.csv("./Data/problems.csv")
problems
```

```{r echo = FALSE}
problems <- read.csv(system.file("extdata", "problems.csv", package = "feedr"))
problems
```

Original bird\_ids:
```{r}
finches$bird_id[1:5]
```

Fix problems and new bird\_ids:
```{r}
r.clean <- check.problems(finches, problems = problems)
r.clean$bird_id[1:5]
```

Note that the bird\_ids have been modified, but also that the levels have been updated to match.

--------------------------
Back to [top](#top)  
Go back to [main document](feedr.html) | Go back to [loading/importing data](load.html) | Continue with [transformations](transformations.html)
