---
title: "Getting started"
---

```{r setup, include = F}
source("setup.R")
```

<!-- Render entire site:
rmarkdown::render_site("./docs/") -->


This is an overview of how to use functions in the `feedr` package to transform and visual your RFID data. Note that while we've tried to explain the code in detail, if you are new to R, you may benefit from an introduction to coding in R before tackling your data with the `feedr` package

## Order of operations
This package contains functions to perform three broad services: 

1. [__Loading/Importing Data__](#load)  
    Raw data directly from the loggers, already amalgamated, or from the animal__nexus__ website 
2. [__Housekeeping__](#hk)  
    Functions for checking and correcting errors before you transform the data
3. [__Transformations__](#trans)  
    Transforming raw data into visits, presence, activity, etc
4. [__Visualizations__](#vis)  
    Static or interactive maps produced by the mapping functions which visualize movements between and presence around the RFID loggers.

There are also three 'user-interface' (UI) functions that can be used to simplify the use of these functions (see below).

In the normal order of operations, several functions depend on the output of other functions.

For example, to get daily activity patterns, you could get your data from the animalnexus database with `dl_data()`, turn this raw data to visits data with `visits()`, turn visits data to activity data with `activity()` and finally, turn activity data into daily activity patterns with `daily()`.

```{r, echo = F, fig.align='center', warning = FALSE}
DiagrammeR::grViz("
digraph order {

graph [overlap = true, compound = true]

node[shape = Mrecord]
edge[arrowhead = vee]

load[label = '{Loading/Importing Data | {load_raw_all() | load_web() | dl_data()}}']

hk[label = '{Housekeeping (Optional) | {check_ids() | check_problems()}}', width = 3]

load -> hk
hk -> visits
load -> visits

visits[label = 'visits()']
disp[label = 'disp()']
dom[label = 'dom()']
move[label = 'move()']
presence[label = 'presence()']
act[label = 'activity()']
daily[label = 'daily()']

visits-> {disp, move, presence}
presence -> act
disp -> dom
act -> daily

map[label = '{Visualizations | {map_ggmap() | map_leaflet()}}']

{presence; move} -> map

#A -> e [ltail = cluster0, lhead = cluster1]
#f -> 'visits()' [ltail = cluster1, lhead = cluster2]
#'move()' -> 'map_leaflet()' [ltail = cluster2, lhead = cluster3]

}
", width = 500)
```

## Tutorial Data
Two data frames, `finches` and `chickadees` are included in the `feedr` package for the purpose of demonstration.

```{r}
head(finches)
head(chickadees)
```

## Function Index

### [Loading/Importing Data](load.html)
<a id = "load"></a>

- __ui_import():__ Launch an interactive user-interface to import/load local data files
- __load_raw():__ Load in a single file of raw logger data
- __load_raw_all():__ Load in and combine multiple files of raw logger data
- __dl_data():__ Download and load data directly from the animal__nexus__ website

### [Housekeeping](housekeeping.html)
<a id = "hk"></a>

- __check_ids():__ Remove known bad ids (i.e. errors, wands, etc.)
- __check_problems():__ Fix known id problems

### [Transformations](transformations.html)
<a id = "trans"></a>

- __ui_trans():__ Launch an interactive user-interface to transform data
- __visits():__ Transform raw data into visits data
- __move():__ Transform visits data into movements between loggers
- __presence():__ Transform visits data into presence around loggers
- __disp():__ Transform visits data into displacements
- __dom():__ Transform displacements into dominance matrices and hierarchies
- __activity():__ Transform visits data into activity data
- __daily():__ Transform activity data into daily activity patterns

## [Visualizations](maps.html)
<a id = "vis"></a>

- __ui_animate():__ Launch an interactive user-interface to create animations
- __map_ggmap():__ Visualize movements and/or presence with ggmap (static)
- __map_leaflet():__ Visualize movements and/or presence with leaflet (interactive)

-----------------------------------------
Back to [top](#top)  
Continue with [loading/importing data](load.html)
