---
title: "Transforming/Summarizing Data"
---

```{r setup, include = F}
source("setup.R")
```

This section will get you started with transforming and summarizing your data from raw read data to a variety of different data types which can be used to answer different questions.

Here we assume that you know how to [load/import](load.html) your data and that, if needed, you've cleaned up problematic `animal_id`s with the various [housekeeping](housekeeping.html) functions.

Note that for several examples of summarizing, we'll be using functions from the [`tidyverse`](http://tidyverse.org/) including the magrittr pipe `%>%`, which essentially 'pipes' the output of the first line to the next line, see this [intro to the pipe](http://magrittr.tidyverse.org/articles/magrittr.html) for more details.

```{r}
library(tidyverse)
```


## Functions

In all cases you'll want to use the `visits()` function first, but after that it depends on what kinda of data you want to end up with.

1. [`visits()`](#visits) - Visits to a logger
2. [`move()`](#movements) - Movements between loggers
3. [`presence()`](#presence) - Presence around a logger
4. [`disp()`](#displacements) - Displacement from a logger
5. [`activity()`](#activity) - Activity patterns
6. [`daily()`](#dailyactivity) - Daily activity patterns
7. [Conversion functions](#conversion-functions) - Functions to convert data for use by other packages
8. [Working with multiple groups](#multiple-groups) - Applying transformations to different data groups simultaneously

Or click on the function name below to skip directly to the function tutorial:

```{r, echo = F, fig.align='center'}
DiagrammeR::grViz("
digraph order {

graph [overlap = true, compound = true]

node[shape = Mrecord]
edge[arrowhead = vee]

visits[label = 'visits()', URL = '#visits', tooltip = 'Individual logger visits']
disp[label = 'disp()', URL = '#displacements', tooltip = 'Displacements from the logger by another animal']
move[label = 'move()', URL = '#movements', tooltip = 'Movements between loggers']
presence[label = 'presence()', URL = '#presence', tooltip = 'presence bouts at loggers']
act[label = 'activity()', URL = '#acitivty', tooltip = 'Activity scored (active/inactive)']
daily[label = 'daily()', URL = '#dailyactivity', tooltip = 'Daily activity patterns']

visits-> {disp, move, presence}
presence -> act
act -> daily

map1[label = 'map_ggmap()', URL = '#mapggmap', tooltip = 'Visualize with a static map']
map2[label = 'map_leaflet()', URL = '#mapleaflet', tooltip = 'Visualize with an interactive map']

{presence; move} -> {map1; map2}

}
", width = 500)
```
<a id = "visits"></a>

## `visits()`
This function is designed to turn 'raw' data into 'visits' data.

This is always the first function you should use. Raw read data contains an individual line of data for each read event made by the RFID logger.

For example, if the logger was set to record ids every ~2 seconds, an individual sitting on the logger for 8 seconds followed by a 12 second gap before returning would look something like this:

```{r}
head(finches_lg)
```

The `visits()` function turns this raw data into something more practical:

```{r visits}
v <- visits(finches_lg)
head(v)
```

We can see that the first series of raw reads have been combined into one visit, starting at 2015-10-01 17:38:52 and ending at 2015-10-01 17:39:00.

We can also look directly at visit data for a particular individual by pulling out it's `animal_id`:

```{r}
head(finches_lg[finches_lg$animal_id == "06200004F8", ])
head(v[v$animal_id == "06200004F8", ])
```

Alternatively, you can use `filter(v, animal_id == "06200004F8")` from the `dplyr` package or `subset(v, animal_id == "06200004F8")` which is built into R base.

Visits are defined by:

- How much time has passed between reads (`bw`)
- A change in `animal_id` between two successive reads
- A change in `logger_id` between two successive reads (although generally this shouldn't be a defining point, because the `bw` should be smaller than the length of time it takes to move from one logger to another)

<a id = "visitsdata"></a>

### Data returned

The data returned by the visit function contains both new and old data:

- `animal_id` id of the animal making the visit
- `date` the date of the start of the visit
- `start` the time of the start of the visit
- `end` the time of the end of the visit
- `logger_id` the id of the logger being visited
- `animal_n` the number of unique `animal_id`s in the data set
- `logger_n` the number of unique `logger_id`s in the data set

Note that several of these new values could also have been obtained by hand:

```{r}
length(unique(v$animal_id))
length(unique(v$logger_id))
```

`species`, `sex`, `lon` and `lat` are all columns originally in the `finches_lg` data set, they were automatically passed through the `visits()` function. Columns of extra data that are associated with either a `date`, an `animal_id`, or a `logger_id` will always be passed through unless `pass = FALSE`:

```{r}
visits(finches_lg, pass = FALSE)
```

### `bw`
Depending on the interval between reads in your RFID logger, you may have to adjust the value of `bw` which specifies how many seconds between reads separates two visits. The default is 3 seconds, but if your loggers scan at a slower interval, you will need to adjust the `bw` argument:

```{r}
visits(finches_lg, bw = 15)
```


### Impossible/missing visits
You should also be aware of 'impossible visits'. These may occur if the internal clocks on your RFIDs are not in sync, or if your loggers are very close to each other. An impossible visit is when an individual animal is detected at two different loggers within 2 seconds (the default) of each other. Unless your loggers are REALLY close to each other, this is highly unlikely.

Because both missing values (NAs) and impossible visits likely indicate a greater underlying problem, you have to specifically tell the function to ignore them (or omit them, in the case of NAs).

```{r eval = FALSE}
v <- visits(finches_lg, allow_imp = TRUE, na_rm = TRUE)
```

You can define the minimum number of seconds needed to travel between loggers to specify which visits should be considered impossible. For example, if we say that it must take animals at least 3 minutes to travel between loggers, any visit made to a different logger within 3 minutes would be defined as 'impossible':

```{r, error = TRUE}
visits(finches_lg, bw_imp = 180)
```

### Summarizing
Visit data is the starting block of all other transformations, but you can also summarize and analyze visit data in itself.

For example, if we want to create daily summaries of the number of visits made and loggers used per `animal_id`, we could use the `dplyr` package (part of the [`tidyverse` family of packages](http://tidyverse.org/)).

Currently we have data on the total number of loggers in the data set, but what if we wanted to know how many loggers each individual used?

```{r}
library(dplyr)
visit_summary <- v %>%
  group_by(animal_id) %>%
  summarize(n_visits = length(start),
            n_loggers = length(unique(logger_id)),
            n_days = length(unique(date)),
            mean_visit_length = mean(as.numeric(difftime(end, start, units = "sec"))))
visit_summary
```

__A Side Note__  
This bit of code uses a slightly advanced technique called 'pipes': `%>%` These allow you to pass the output from one line directly down to the next line where it becomes the first (and hidden) argument. In this case, the first line specifies our visits data frame `v`, this then becomes the input for `group_by()`. The output of `group_by()` then becomes the input for `summarize()`. This makes your coding more efficient but may take a bit of getting used to.

You can find an in-depth look at `dplyr`, pipes and related packages here: <http://r4ds.had.co.nz/transform.html>

For reference, you can achieve the same effect without pipes:

```{r}
library(dplyr)
visit_summary <- group_by(v, animal_id)
visit_summary <- summarize(visit_summary,
               n_visits = length(start),
               n_loggers = length(unique(logger_id)),
               n_days = length(unique(date)),
               mean_visit_length = mean(as.numeric(difftime(end, start, units = "sec"))))
visit_summary
```

The result is a data frame with one row for each `animal_id`, giving us information on how many loggers that animal has used, how many visits they have made, over how many days, and what the average visit length was.

If we wanted to plot this data, the package `ggplot2` is a great option:

```{r, fig.width = 6, fig.height = 3}
library(ggplot2)
ggplot(data = visit_summary, aes(x = animal_id, y = mean_visit_length)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + #rotate/shift id labels
  geom_bar(stat = "identity") + # Create bars
  labs(x = "Animal ID", y = "Average visit length (s)") # Make labels pretty

```

--------------------
Back to [top](#top) 
--------------------

<a id = "movements"></a>

## `move()`

This function is designed to turn 'Visit' data into 'Movement' data.

**Movements** are defined as trips between loggers (i.e. a event when a visit from one `animal_id` occurs at a different logger from the previous visit by that particular `animal_id`).

```{r move}
v <- visits(finches_lg, bw = 5)
m <- move(v)
head(m)
```

### Data returned

This function returns a lot of information for each movement event. Note, in particular, that each movement event results in two rows of data: leaving and arriving

- `animal_id` id of the animal making the move
- `date` the date of the event (either arriving or leaving)
- `time` the time of the event (either arriving or leaving)
- `logger_id` the logger left or arrived at
- `direction` leaving or arriving
- `move_id` for each individual, a unique number tying the two events (leaving/arriving) of the movement together
- `move_dir` category indicating direction of the movement and the loggers involved
- `move_path` category indicating the loggers involved (no directionality)
- `strength` a metric indicating how 'easily' individuals move been loggers, it represents 1/time between loggers

`animal_n`, `logger_n`, `species`, `sex`, `lon`, and `lat` are extra columns passed through (see [visits data](#visitsdata) for more information).


### Include all

In this case, four `animal_id` have been omitted because they never moved between loggers. While the id is retained in the _levels_ of the factor, `animal_id`, it is not present in the data frame:

```{r}
unique(m$animal_id)         # The animal_ids currently in the movement data
length(unique(m$animal_id)) # The number of animal_ids currently in the data
length(levels(m$animal_id)) # The number of levels of animal_id
```

However, if you wanted to retain all animals in the data set, regardless of whether they moved or not (for example, if you wanted to look at the likelihood of moving between loggers), you can specify that:

```{r}
m <- move(v, all = TRUE)

unique(m$animal_id)         # The animal_ids currently in the movement data
length(unique(m$animal_id)) # The number of animal_ids currently in the data
length(levels(m$animal_id)) # The number of levels of animal_id
```

Here we've told R that an additional parameter for `move()` is `all = TRUE`. animals which never visit more than one logger will have NA's in the data frame:

```{r}
m[is.na(m$direction), ]
```

### Summarizing

Once again we can summarize movement data using the `dplyr` package.

Are certain paths used more often?
```{r}
move_summary <- m %>%
  group_by(move_path, logger_id) %>%
  summarize(n = length(move_path))
move_summary
```
The NAs reflect the four individuals that did not move. 

Note that because there are two rows per movement event (leaving and arriving), we group by logger_id to summarize individually by logger (resulting in duplicate `n`s per `move_path`. This will account for the 2 rows and keep logger ids in the final data frame (required for mapping later on).

However, if you only wanted `move_path` and `n`, you could simply omit `logger_id` and remove duplicates:
```{r}
move_summary %>% 
  select(-logger_id) %>%
  unique()
```

```{r, fig.width = 6}
ggplot(data = move_summary, aes(x = move_path, y = n)) +
  geom_bar(stat = "identity")
```

Are certain directions on particular paths used more often? (i.e. do birds go back and forth or do they circle around?)
```{r}

move_summary <- m %>%
  group_by(logger_id, move_path, move_dir) %>%
  summarize(n = length(move_path))
head(move_summary)
```

```{r, fig.width = 6}
ggplot(data = move_summary, aes(x = move_path, y = n, fill = move_dir)) +
  geom_bar(stat = "identity", position = "dodge", colour = "black")
```

Is there a relationship between use of a path and the average 'strength'?
```{r}

move_summary <- m %>%
  group_by(logger_id, move_path) %>%
  summarize(n = length(move_path),
            strength = mean(strength, na.rm = TRUE))
move_summary
```

```{r}
ggplot(data = move_summary, aes(x = strength, y = n)) +
  geom_point() +
  stat_smooth(method = "lm") # add a regression line w/ standard error
```

Do some animals move more than others?
```{r}
move_summary <- m %>%
  group_by(animal_id) %>%
  summarize(n = length(move_path[!is.na(move_path)])) # Number of movements omitting NAs
move_summary
```

Note that we have some zeros because we made sure to include all individuals in the analysis (`all = TRUE`, above) and because we specified `summarize()` to count only non-NA paths (`move_path[!is.na(move_path)]`).

```{r, fig.width = 6}
ggplot(data = move_summary, aes(x = animal_id, y = n)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + #rotate and shift id labels
  geom_bar(stat = "identity") 
```


--------------------
Back to [top](#top) 
--------------------

<a id = "presence"></a>

## `presence()`

This function is designed to turn 'Visit' data into 'Presence' data.

**Presence** reflects blocks of time in which an animal visited a given logger regularly. Essentially, presence is defined as a series of visits, and can overlap with presence of other animals (this differs from [visits](#visits), as normally one individual's visits cannot overlap with another). Presence near a logger ends when the individual moves to another logger or when a particular amount of time has passed without a visit (the time between visits, `bw`, 15 min by default).

```{r presence}
v <- visits(finches_lg, bw = 5)
p <- presence(v)
head(p)
```

### Data returned
- `animal_id` id of the animal present
- `date` the date of the start of the event
- `logger_id` the logger involved
- `start` the start time of the event
- `end` the end time of the event
- `length` the length of the event in minutes

`animal_n`, `logger_n`, `species`, `sex`, `lon`, and `lat` are extra columns passed through (see [visits data](#visitsdata) for more information).


### `bw`
The user of the function may define what the minimum amount of time between visits should be. Ground truthing could help establish the amount of time an individual spends near a logger without making a logger visit and could be used to determine this cutoff time. 

Note that the units for `bw` when defining [visits](#visits) is in seconds, while the units for `bw` when defining presence is in minutes. This reflects the different scales of the two functions.

```{r}
v <- visits(finches_lg, bw = 5)
p <- presence(v, bw = 20)  # Within 20 min
head(p)
```

```{r}
v <- visits(finches_lg, bw = 5)
p <- presence(v, bw = 90)  # Within 1.5 h
head(p)
```

The `bw` parameter can also be omitted entirely:
```{r}
v <- visits(finches_lg, bw = 5)
p <- presence(v, bw = NULL)
head(p)
```

When `bw = NULL` only a move to a different logger will initiate a new presence bout. Therefore this identifies how much time passes before a new logger is visited. The end time reflects the time of the final visit to that logger before moving to a new logger, but if an individual were to visit logger A all day on day 1 and disappear on day 2 and then make one visit to logger A on day 3 before moving immediately to logger B, that first bout of 'presence' would be considered two days long, whereas in reality, the individual did not make any visits on one of those days.

### Summarizing

What's the average time present for each individual?
```{r}
v <- visits(finches_lg, bw = 5)
p <- presence(v)

presence_summary <- p %>%
  group_by(animal_id) %>%
  summarize(n = length(start),          ## how many rows = number of events
            mean_length = mean(length)) ## mean length of events

presence_summary
```

Is there a relationship between the number and length of presence bouts?
```{r}
ggplot(data = presence_summary, aes(x = n, y = mean_length)) +
  geom_point()
```

Do animals spend more time around certain loggers?
```{r}
presence_summary <- p %>%
  group_by(logger_id) %>%
  summarize(n = length(start),
            mean_length = mean(length),
            se_length = sd(length)/sqrt(n))

presence_summary

ggplot(data = presence_summary, aes(x = logger_id, y = mean_length, fill = logger_id)) +
  geom_bar(stat = "identity") +
  geom_errorbar(aes(ymin = mean_length - se_length, 
                    ymax = mean_length + se_length), width = 0.25)
```

Note that there is only one bout of presence around logger 2400, hence the lack of errorbars.

--------------------
Back to [top](#top) 
--------------------

<a id = "displacements"></a>

## `disp()`
This function is designed to turn 'visit' data into 'displacements' data.

__Displacements__ are events when one animal leaves the logger and is replaced by another animal within a certain, defined, time span.

In theory, we assume this is due to the arrival of a more dominant animal. Displacements events might therefore lead to information on relative dominance of different individuals.

> __Disclaimer!__  
> It is important to ground-truth this assumption because, depending on species, displacements may not necessarily represent dominance. For example, in the case of black-capped chickadees, if the dominant male is using the logger, subsequent users of the logger may actually arrive and depart in descending order of rank. In this case, apparent 'displacements' may not be displacements at all.

```{r disp}
v <- visits(finches_lg, bw = 5)
d <- disp(v)
names(d)
```

### Data returned
This function returns a list with three different types of displacement information: `displacements`, `summaries`, and `interactions`.

```{r}
head(d$displacements)
```

The `displacements` item is a data frame containing information on each displacement event

- `animal_id` id of an animal involved in a displacement event
- `date` the date of the start of the event
- `left` the time the 'displacee' left the logger
- `arrrived` the time the 'displacer' arrived at the logger
- `logger_id` the logger involved in the event
- `role` the role in the event of the current animal ('displacee' or 'displacer')

`animal_n`, `logger_n`, `species`, `sex`, `lon`, and `lat` are extra columns passed through (see [visits data](#visitsdata) for more information).


```{r}
head(d$summaries)
```

The `summaries` item is a data frame summarizing the displacement events for each individual

- `animal_id` id of animal
- `displacee` is the number of times the animal was the displacee
- `displacer` is the number of times the animal was the displacer
- `p_win` is the proportion of times the animal 'won' the interaction (i.e. number of displacer events divided by total interactions)

```{r}
head(d$interactions)
```

The `interactions` item is a data frame summarizing each potential interaction event:

- `displacer` is id of the animal doing the displacement
- `displacee` is id of the animal being displaced
- `n` is the number of times that particular combination of displacer and displacee was observed

### No events

If there are no displacement events in your data set, the function will give you a message and will return an empty list. In this case, the data set is so small, there are no displacement events:

```{r}
v <- visits(finches, bw = 5)
disp(v)
```

### `bw`

The `disp()` function also contains a `bw` argument. In this case it reflects the maximum interval between the first animal leaving and the second animal arriving for an interaction to be considered a displacement:

```{r}
v <- visits(finches_lg, bw = 5)
head(disp(v)$summaries)
head(disp(v, bw = 10)$summaries)
head(disp(v, bw = 15)$summaries)
```
However, an interval of 15s is pretty long and probably doesn't constitute a real displacement event.

### Summarizing

Is one logger more contentious than others?

```{r}
v <- visits(finches_lg, bw = 5)
d <- disp(v)

disp_summary <- d$displacements %>%
  group_by(logger_id, role) %>%
  count() %>%
  filter(role == "displacer")

ggplot(data = disp_summary, aes(x = logger_id, y = n)) +
  geom_bar(stat = "identity")
```

What if we control for overall logger use?

```{r}
disp_summary <- d$displacements %>%
  group_by(logger_id, role) %>%
  count() %>%
  filter(role == "displacer")

presence_summary <- p %>%
  group_by(logger_id) %>%
  summarize(amount = sum(as.numeric(length)))

full_summary <- full_join(disp_summary, presence_summary, by = "logger_id") %>%
  mutate(n = replace(n, is.na(n), 0),  # First replace the missing values of n with 0
         n_per_hour = n / amount * 60) # Calculate the number of interactions as an hourly rate

full_summary

ggplot(data = full_summary, aes(x = logger_id, y = n)) +
  geom_bar(stat = "identity")
```


```{r}
d$summaries

ggplot(data = d$summaries, aes(x = animal_id, y = p_win)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + #rotate and shift id labels
  geom_bar(stat = "identity") +
  ylim(c(0, 1)) +
  labs(x = "Animal Id", y = "Proportion of 'wins'")

```

--------------------
Back to [top](#top) 
--------------------

<a id = "activity"></a>

## `activity()`
This function scores individuals as either active or inactive based on their presence data. While simply looking at the presence bouts is a good proxy for active/inactive, scoring intervals as active/inactive allows for easier plotting as well as for averaging over days to create daily activity patterns (see [`daily()` function](#dailyactivity)).

```{r activity}
v <- visits(finches_lg, bw = 5)
p <- presence(v)
a <- activity(p, res = 5)
head(a)
```

### Data returned

Several messages along with a data set containing:

- `animal_id` id of the animal in question
- `date` the date of the event
- `time` the time block of the event
- `activity` a numeric activity score: 0 (inactive), 1 (active)
- `activity_c` a categorical activity score: inactive, active
- `logger_id` the logger where the activity occurred. If summarizing across loggers, this will be `NA`.
- `rise` the time of sunrise (median time if calculated across all loggers), only calculated if `sun = TRUE` 
- `set` the time of sunset (median time if calculated across all loggers), only calculated if `sun = TRUE` 

`animal_n`, `logger_n`, `species`, `sex` are extra columns passed through (see [visits data](#visitsdata) for more information).

The messages reflect information regarding the quality of the activity data:

- activity being low ("Active less than 5%...") 
- skipped individuals ("Skipping. Individual has less than 24hrs data) (see `keep_all` below)
- obs which are shorter than the resolution 'res' (X% of obs are shorter than 'res') (see `res` below)


### `res` 
The default function calculates activity for each animal at a resolution of 15min. Meaning that every 15min an individual is scored as active or inactive. However, if individuals make short and infrequency visits to a logger, periods of activity may be missed if the active bouts are smaller than the resolution. Resolution can be adjusted through the `res` argument (note the change in time, and the change in messages).

```{r, message = FALSE}
head(activity(p, res = 1))
```

### `by_logger`
By default, the `activity()` function sums all activity across loggers. However, if you wish to calculate separate activity patterns for different loggers, you can set `by_logger = TRUE`. Note the `logger_id` and that there are now 4 observations per time block for each individual:

```{r, message = FALSE}
head(activity(p, by_logger = TRUE))
```

### `sun`
By default sunrise and sunset times are calculated if the data contains latitude and longitude information. You can turn this off with `sun = FALSE`.

### `keep_all`
By default, the `activity()` function skips individuals with less than 24hrs of data, you can force it to retain all individuals by specifying `keep_all = TRUE`.

### `missing`
Currently this functionality is not implemented. However, eventually we plan to allow users to specify times of known missing data, so that activity can be specified as 'active', 'inactive', or 'unknown'.


### Summarizing

Activity patterns can be plotted in ggplot2. Let's look at a specific time period:

```{r, fig.height = 10}
library(dplyr)
library(ggplot2)
library(lubridate)

# Filter by date and animal ID, create a "day" column
sept <- a %>%
  filter(date >= as.Date("2015-09-21"),
         date < as.Date("2015-09-28"),
         animal_id == "06200004F8")

# Use facet_wrap to look at individual days, use drop = TRUE to omit days without activity
ggplot(data = sept, aes(x = time, y = activity)) +
  facet_wrap(~ date, scales = "free_x", ncol = 1, drop = TRUE) +
  geom_area()
```

There's a clear daily pattern, but it's a bit hard to tell without sunrise and sunset:

```{r, fig.height = 10}
sun <- unique(select(sept, rise, set, date))

ggplot(data = sept, aes(x = time, y = activity)) +
  facet_wrap(~ date, scales = "free_x", ncol = 1, drop = TRUE) +
  geom_rect(mapping = aes(xmax = rise), xmin = -Inf, ymin = -Inf, ymax = +Inf, fill = "grey") +
  geom_rect(mapping = aes(xmin = set), xmax = +Inf, ymin = -Inf, ymax = +Inf, fill = "grey") +
  geom_area()
```

--------------------
Back to [top](#top) 
--------------------

<a id = "dailyactivity"></a>

## `daily()`

`daily()` converts activity patterns created by the [`activity()` function](#activity) to averaged daily activity patterns.

```{r daily, message = FALSE}
v <- visits(finches_lg, bw = 5)
p <- presence(v)
a <- activity(p, res = 5)
da <- daily(a)
```

### Data returned
A data set containing:

- `animal_id` id of the animal in question
- `time` the time block of the event (note the odd date)
- `time_c` the time block of the event (categorical)
- `p_active` the proportion of times this block was scored as active
- `p_inactive` the proportion of times this block was scored as inactive
- `p_unknown` the proportion of times this block was scored as unknown
- `p_total` the proportion of times this block was scored as anything but unknown (i.e., 1 - `p_unknown`)
- `logger_id` the logger where the activity occurred. If summarizing across loggers, this will be `NA`.
- `rise` the time of sunrise (median time if calculated across all loggers), only calculated if `sun = TRUE` 
- `set` the time of sunset (median time if calculated across all loggers), only calculated if `sun = TRUE` 

`animal_n`, `logger_n`, `species`, `sex` are extra columns passed through (see [visits data](#visitsdata) for more information).

If missing data are indicated in `activity()`, then the proportion of unknown blocks will be greater than 0 (Note that this is not currently available).

### Summarizing
Daily activity patterns can be plotted by hiding the date portion of the time stamp.

```{r}
ggplot(data = da, aes(x = time, y = p_active)) +
  geom_rect(aes(xmax = rise), xmin = -Inf, ymin = -Inf, ymax = +Inf, fill = "grey") +
  geom_rect(aes(xmin = set), xmax = +Inf, ymin = -Inf, ymax = +Inf, fill = "grey") +
  geom_area() +
  facet_wrap(~ animal_id) +
  scale_x_datetime(date_labels = "%H:%M:%S", date_breaks = "6 hours")
```

Note also that this script repeats the drawing of the sunrise/sunset rectangles, making it difficult to use alpha values. To avoid this, create a reduced data frame with only sunrise/sunset data:

```{r}
sun <- unique(da[, c("rise", "set", "animal_id")])

ggplot(data = da) +
  geom_rect(data = sun, aes(xmax = rise), xmin = -Inf, ymin = -Inf, ymax = +Inf, fill = "black", alpha = 0.1) +
  geom_rect(data = sun, aes(xmin = set), xmax = +Inf, ymin = -Inf, ymax = +Inf, fill = "black", alpha = 0.1) +
  geom_area(aes(x = time, y = p_active)) +
  facet_wrap(~ animal_id) +
  scale_x_datetime(date_labels = "%H:%M:%S", date_breaks = "6 hours")
```

## Conversion functions

Conversion functions permit you to format raw or transformed data for use in other packages.

### `convert_anidom()`

Converts the output of `disp()` for use by the package [`aniDom`](https://cran.r-project.org/package=aniDom)'s functions: `elo_scores()` and `estimate_uncertainty_by_splitting()`

```{r}
d <- disp(visits(finches_lg), bw = 5)
i <- convert_anidom(d)
head(i)
```

```{r}
library(aniDom)

# Calculate elo_scores
s <- elo_scores(winners = i$winner, losers = i$loser)
s

# Estimate repeatability
estimate_uncertainty_by_repeatability(winners = i$winner, losers = i$loser)
estimate_uncertainty_by_splitting(winners = i$winner, losers = i$loser, 
                                  randomise = TRUE)
```

### `convert_dominance()`
Converts the output of `disp()` for use by the package [`Dominance`](https://cran.r-project.org/package=Dominance)'s functions: `ADI()` and `Sociogram()`

```{r}
d <- disp(visits(finches_lg), bw = 5)
i <- convert_dominance(d)

library(Dominance)

# Calculate the Average Dominance Index
ADI(data_sheet = i$data_sheet, items = i$items, actions = i$actions, bytes = i$bytes)
```

```{r eval = FALSE}
# Construct social network graphs
Sociogram(data_sheet = i$data_sheet, items = i$items, actions = i$actions, bits = i$bytes)
```

### `convert_perc()`
Converts the output of `disp()` for use by the package [`Perc`](https://cran.r-project.org/package=Perc)

```{r}
d <- disp(visits(finches_lg), bw = 5)
i <- convert_perc(d)
head(i)
```

```{r}
library(Perc)

# Calculate ranks (adapted from Perc examples)
conflict_mat <- as.conflictmat(i, weighted = TRUE)
perm <- conductance(conflict_mat, 2)
simRankOrder(perm$p.hat, num = 10, kmax = 1000)
```

### `convert_asnipe()`
Converts raw RFID data into a format for easy use by either the `gmmevents()` or the `get_associations_points_tw()` functions included in the [`asnipe`](https://cran.r-project.org/package=asnipe) package for calculating group membership.

For `gmmevents()`:
```{r}
a <- convert_asnipe(finches, fun = "gmmevents")
head(a)
```

```{r, eval = FALSE}
library(asnipe)
gmmevents(a$time, a$identity, a$location)
```

For `get_associations_points_bw()`:
```{r}
a <- convert_asnipe(finches, fun = "get_associations_points_tw")
head(a)
```

```{r, eval = FALSE}
get_associations_points_tw(a)
```


### `convert_activity()`
Converts raw RFID data into a format for easy use by the `fitact()` function included in the [`activity`](https://cran.r-project.org/package=activity) package for modelling activity levels and daily activity patterns.

```{r}
i <- convert_activity(finches_lg)
```

```{r}
library(activity)

# Calculate daily activity pattern for a single individual
a <- fitact(i[[1]], sample = "none")
plot(a)

# Calculate daily activity pattern for all individuals
a <- lapply(i, fitact, sample = "none")
plot(a[[3]])
plot(a[["06200004F8"]])
```


--------------------
Back to [top](#top) 
--------------------


## Multiple groups

Sometimes you may want to perform the same data transformations multiple times, perhaps for different sites, time periods, or experiments. You could repeat yourself, but the easiest thing to do is to use the `do()` function from the [`dplyr` package](http://dplyr.tidyverse.org/) or the `map()` frunction from the [`purrr` package](http://purrr.tidyverse.org/).

These function allows you to apply functions to groups in your data frame.

In this example we have two experiments, so we want to calculate the transformations separately for each experiment:

```{r}
library(dplyr)

v_exp <- chickadees %>%
  group_by(experiment) %>%
  do(visits(.))  # Use "." to define where the piped data should go 

# Note that the data set retains it's grouping so we do not have to 
# specify the grouping again until we change it:
p_exp <- v_exp %>%
  do(presence(.)) 

m_exp <- v_exp %>%
  do(move(.))

# Summarizing
p_avg <- p_exp %>%
  group_by(experiment, logger_id) %>%
  summarize(amount = sum(length)/animal_n[1])  #This is using the animal_n specific to the experiment

m_avg <- m_exp %>%
  group_by(experiment, move_path, logger_id, lat, lon) %>%
  summarize(path_use = length(move_path))

summary(p_avg)
summary(m_avg)
```

These data are now ready for analysis and visualizations, note that they can be applied as is to visualizing functions in the next section.


In future versions of `dplyr` it is likely the `do()` function will be deprecated, so let's consider the `map()` function from the `purrr` package. This is also an interesting method in that all data sets can be kept together in a single object (note that each line in mutate runs sequentially, so we can create the column `visits` and then use it in the next line).

```{r}
library(dplyr)
library(purrr)

exp <- chickadees %>%
  group_by(experiment) %>%
  nest(.key = "raw") %>%
  mutate(visits = map(raw, ~ visits(.x)),
         presence = map(visits, ~ presence(.x)),
         move = map(presence, ~ move(.x)))
         
exp
```


## Summarizing

```{r}

p_avg <- exp %>%
  unnest(presence) %>%
  group_by(experiment, logger_id) %>%
  summarize(amount = sum(length)/animal_n[1])  
  # This is using the animal_n specific to the experiment

m_avg <- exp %>%
  unnest(move) %>%
  group_by(experiment, move_path, logger_id, lat, lon) %>%
  summarize(path_use = length(move_path))

summary(p_avg)
summary(m_avg)
```


--------------------------
Back to [top](#top)  
Go back to [home](index.html) | Go back to [housekeeping](housekeeping.html) | Continue with [visualizations](maps.html)
